---
title: "BIOSTAT675 - hw1"
author: "Nianlin Chen"
date: "9/22/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r}
df <-read.table("/Users/cassiechen/Downloads/Breast_cancer_Table_1_2_Collet.txt")
library(survival)
```

## a)Compute the Nelson-Aalen estimator of H(t)
```{r}
fit <- survfit(Surv(V2,V3)~1,data=df,conf.type="plain" )
h_j <- fit$n.event/fit$n.risk
Ht_NA <- cumsum(h_j)
t_j <- fit$time
d_j <- fit$n.event
Y_j <- fit$n.risk
df1  <- data.frame(tj = t_j, dj = d_j, Yj = Y_j, hj = h_j,Ht_NA =Ht_NA)
df1
```

## b) Compute 95% CI for H(t). Assume hat H(t) follows a normal distribution.
```{r}
sd_Ht= sqrt(cumsum(fit$n.event/(fit$n.risk)^2))
Upper_CI_1 <- Ht_NA + 1.96*sd_Ht
Lower_CI_1 <- Ht_NA - 1.96*sd_Ht
CI_Ht_normal <-data.frame(sd_Ht= sd_Ht,Upper_CI = Upper_CI_1,Lower_CI = Lower_CI_1)
CI_Ht_normal
```
## c) Compute 95% CI for H(t). Assume log hat H(t) follows a normal distribution.
```{r}
var_ht = cumsum(fit$n.event/(fit$n.risk)^2)
sd_log_Ht = sqrt((1/(df1$Ht_NA)^2)*var_ht)
Upper_CI_2 <- Ht_NA + exp(1.96*sd_log_Ht)
Lower_CI_2 <- Ht_NA - exp(1.96*sd_log_Ht)
CI_Ht_log <-data.frame(sd_log_Ht = sd_log_Ht,Lower_CI = Lower_CI_2,Upper_CI = Upper_CI_2)
CI_Ht_log
```

## d) Estimate S(t) using the Nelson -Aalen prodecure, then compute 95% confidence intervals for and assuming that log(hat_H(t)) follows the normal distribution.
```{r}
df1$S_NA_t <- exp(-Ht_NA)
Upper_CI_3 = df1$S_NA_t + exp(-exp(1.96*sd_log_Ht))
Lower_CI_3 = df1$S_NA_t - exp(-exp(1.96*sd_log_Ht))
CI_St_log = data.frame(S_NA_t=df1$S_NA_t, sd_log_Ht=sd_log_Ht,Lower_CI = Lower_CI_3, Upper_CI = Upper_CI_3)
CI_St_log
```

## e) etimate the median and quartiles of the survival distribution.
```{r}
quantile(df1$S_NA_t)
```

## f)  Estimate mean residual life,E[T−t|T > t], truncated at 15 years, at all observed death times prior to 15 years.
t0 = 5 years, truncated at 15 years, which is t =180
```{r}
index = which(df1$tj <= 180)
inter_i <- 0
for (i in 1:length(index)){
  inter_i[i] <- (df1$tj[i+1] - df1$tj[i])* df1$S_NA_t[i]
}
sum(inter_i)/df1$S_NA_t[index[1]]
```